Tree height and hydraulic traits shape growth responses across droughts in a temperate broadleaf forest

### 2020-09-22 manuscript/climate_sensitivity_draft.Rmd
Updated to make sure citations worked #147

---
title: ""
author: ""
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
bibliography: ["book.bib",  "packages.bib"]
link-citations: yes
biblio-style: "apalike"
csl: new-phytologist.csl
header-includes:
- \usepackage{setspace}\onehalfspacing
- \usepackage{lineno}
- \linenumbers
- \usepackage{float}
- \usepackage{booktabs}
- \usepackage{pdflscape}
- \newcommand{\blandscape}{\begin{landscape}}
- \newcommand{\elandscape}{\end{landscape}}
- \usepackage{caption}
- \captionsetup[table]{font=small}
- \captionsetup[figure]{font=small}
- \captionsetup[table]{labelformat=empty}
- \captionsetup[figure]{labelformat=empty}
- \usepackage{dcolumn} #for sidewaystable
---

@@

Research was conducted at the 25.6-ha ForestGEO (Forest Global Earth Observatory) study plot at the Smithsonian Conservation Biology Institute (SCBI) in Virginia, USA (38\°53'36.6"N, 78\°08'43.4"W, elevation 273-338 m.a.s.l.; Fig. **S1**) [@bourg_initial_2013; @andersonteixeira_ctfs-forestgeo:_2015]. Climate is humid temperate, with mean annual temperature of 12.7^$\circ$^C and precipitation of 1005 mm yr^-1^ during our study period [1960-2009; source: CRU TS v.4.01; @harris_updated_2014]. Dominant tree taxa within this secondary forest include *Liriodendron tulipifera*,  oaks (*Quercus* spp.), and hickories (*Carya* spp.; Table 2).

@@

As climate change drives increasing drought in many of the world's forests [@trenberth_global_2014; @intergovernmental_panel_on_climate_change_climate_2015], the fate of forests and their climate feedbacks will be shaped by the biophysical and physiological drivers observed here. 
Our results show that taller, more exposed trees and species with less drought-tolerant leaf traits will be most affected in terms of both growth during the drought year and subsequent growth.
Survival is linked to resistance and resilience [@desoto_low_2020; @gessler_way_2020], implying it may be influenced by the same factors. 
Indeed, while no link between $PLA_{dry}$ or $\pi_{tlp}$ on drought survival has been established [but see @powers_catastrphic_2020], taller trees have lower survival [@bennett_larger_2015; @stovall_tree_2019]. 
As climate change-driven droughts affect forests worldwide, there is likely to be a shift from mature forests with tall, buffering trees to forests with a shorter overall stature [@mcdowell_pervasive_2020]. 
At this point, species whose drought tolerance relies in part on existence within a buffered microenvironment (e.g., *F. grandifolia*) could in turn become more susceptible. 
Here, the relative importance of tree height *per se* versus crown exposure becomes crucial, shaping whether the dominant trees of shorter canopies are significantly more drought tolerant because of their shorter stature, or whether high exposure makes them as vulnerable as the taller trees of the former canopy.
Studies disentangling the influence of height and exposure on drought tolerance will be critical to answering this question. 
Ultimately, distributions of tree heights and drought tolerance traits across broad moisture gradients suggest that forests exposed to more drought will shift towards shorter stature and be dominated by species with more drought-tolerant traits [@liu_hydraulic_2019; @bartlett_correlations_2016; @zhu_leaf_2018].
Our study helps to elucidate the mechanisms behind these patterns, opening the door for more accurate forecasting of forest responses to future drought.

>>>>

---
title: ""
author: ""
output:
  pdf_document:
    fig_caption: yes
bibliography: ["book.bib",  "packages.bib"]
link-citations: yes
biblio-style: "apalike"
csl: new-phytologist.csl
header-includes:
- \usepackage{setspace}\onehalfspacing
- \usepackage{lineno}
- \linenumbers
- \usepackage{float}
- \usepackage{booktabs}
- \usepackage{pdflscape}
- \newcommand{\blandscape}{\begin{landscape}}
- \newcommand{\elandscape}{\end{landscape}}
- \usepackage{caption}
- \captionsetup[table]{font=small}
- \captionsetup[figure]{font=small}
- \captionsetup[table]{labelformat=empty}
- \captionsetup[figure]{labelformat=empty}
- \usepackage{dcolumn} #for sidewaystable
---

@@

Research was conducted at the 25.6-ha ForestGEO (Forest Global Earth Observatory) study plot at the Smithsonian Conservation Biology Institute (SCBI) in Virginia, USA (38\°53'36.6"N, 78\°08'43.4"W, elevation 273-338 m.a.s.l.; Fig. **S1**) [@bourg_initial_2013; @andersonteixeira_ctfs-forestgeo_2015]. Climate is humid temperate, with mean annual temperature of 12.7^$\circ$^C and precipitation of 1005 mm yr^-1^ during our study period [1960-2009; source: CRU TS v.4.01; @harris_updated_2014]. Dominant tree taxa within this secondary forest include *Liriodendron tulipifera*,  oaks (*Quercus* spp.), and hickories (*Carya* spp.; Table 2).

@@

As climate change drives increasing drought in many of the world's forests [@trenberth_global_2014; @intergovernmental_panel_on_climate_change_climate_2015], the fate of forests and their climate feedbacks will be shaped by the biophysical and physiological drivers observed here. 
Our results show that taller, more exposed trees and species with less drought-tolerant leaf traits will be most affected in terms of both growth during the drought year and subsequent growth.
Survival is linked to resistance and resilience [@desoto_low_2020; @gessler_way_2020], implying it may be influenced by the same factors. 
Indeed, while no link between $PLA_{dry}$ or $\pi_{tlp}$ on drought survival has been established [but see @powers_catastrophic_2020], taller trees have lower survival [@bennett_larger_2015; @stovall_tree_2019]. 
As climate change-driven droughts affect forests worldwide, there is likely to be a shift from mature forests with tall, buffering trees to forests with a shorter overall stature [@mcdowell_pervasive_2020]. 
At this point, species whose drought tolerance relies in part on existence within a buffered microenvironment (e.g., *F. grandifolia*) could in turn become more susceptible. 
Here, the relative importance of tree height *per se* versus crown exposure becomes crucial, shaping whether the dominant trees of shorter canopies are significantly more drought tolerant because of their shorter stature, or whether high exposure makes them as vulnerable as the taller trees of the former canopy.
Studies disentangling the influence of height and exposure on drought tolerance will be critical to answering this question. 
Ultimately, distributions of tree heights and drought tolerance traits across broad moisture gradients suggest that forests exposed to more drought will shift towards shorter stature and be dominated by species with more drought-tolerant traits [@liu_hydraulic_2019; @bartlett_correlations_2016; @zhu_leaf_2018].
Our study helps to elucidate the mechanisms behind these patterns, opening the door for more accurate forecasting of forest responses to future drought.
